---
title: "APAN5420 --- HW 1"
author: 'Megan Wilder'
date: "5/25/18"
output:
  pdf_document: 
    number_sections: yes
    toc: yes
    toc_depth: 1
  html_document: 
    df_print: default
    number_sections: yes
    toc: no
    toc_depth: 1
---

-----

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

#Load both ETF and Stock Data
```{r}
#load packages
library(readr)  # for read_csv()
library(dplyr)  # for mutate()
library(tidyr)  # for unnest()
library(purrr)  # for map(), reduce()
library(stringr) # extract symbols

# find all file names ending in .txt
files <- dir(pattern = "*.txt")

#load ETF and Stock data and create symbol column with ticker from file name

DF = tibble(File = files) %>%
  extract(File, "Symbol", "([^.]+)", remove = FALSE) %>%
  mutate(Data = lapply(File, read_csv)) %>%
  unnest(Data) %>%
  select(-File)

```

#Explore Data
```{r}
#Summary
summary(DF)

#View NA's in DF
new_DF_open <- DF[is.na(DF$Open),]
new_DF_high <- DF[is.na(DF$High),]
new_DF_low <- DF[is.na(DF$Low),]
new_DF_close <- DF[is.na(DF$Close),]
new_DF_volume <- DF[is.na(DF$Volume),]

unique(new_DF_open$Symbol)
unique(new_DF_high$Symbol)
unique(new_DF_low$Symbol)
unique(new_DF_close$Symbol)
unique(new_DF_volume$Symbol)
#tvix and uvxy, aezs, brk_a, brk_b, clbs, fbc, rgse and bac did not load properly

#remove tvix and uvxy, aezs, brk_a, brk_b, clbs, fbc, rgse and bac data and reload
DF = DF[!DF$Symbol == "tvix", ]
DF = DF[!DF$Symbol == "uvxy", ]
DF = DF[!DF$Symbol == "aezs", ]
DF = DF[!DF$Symbol == "brk-a", ]
DF = DF[!DF$Symbol == "brk-b", ]
DF = DF[!DF$Symbol == "clbs", ]
DF = DF[!DF$Symbol == "fbc", ]
DF = DF[!DF$Symbol == "rgse", ]
DF = DF[!DF$Symbol == "bac", ]

#reload tvix and uvxy, aezs, brk_a, brk_b, clbs, fbc, rgse and bac
tvix <- read.csv("tvix.us.txt")
uvxy <- read.csv("uvxy.us.txt")
aezs <- read.csv("aezs.us.txt")
brk_a <- read.csv("brk-a.us.txt")
brk_b <- read.csv("brk-b.us.txt")
clbs <- read.csv("clbs.us.txt")
fbc <- read.csv("fbc.us.txt")
rgse <- read.csv("rgse.us.txt")
bac <- read.csv("bac.us.txt")

#add symbol column
tvix$Symbol <- "tvix"
uvxy$Symbol <- "uvxy"
aezs$Symbol <- "aezs"
brk_a$Symbol <- "brk_a"
brk_b$Symbol <- "brk_b"
clbs$Symbol <- "clbs"
fbc$Symbol <- "fbc"
rgse$Symbol <- "rgse"
bac$Symbol <- "bac"

#attach tvix and uvxy, aezs, brk_a, brk_b, clbs, fbc, rgse and bac to DF_new
DF_new <- rbind(DF, aezs)
DF_new <- rbind(DF_new, brk_a)
DF_new <- rbind(DF_new, brk_b)
DF_new <- rbind(DF_new, clbs)
DF_new <- rbind(DF_new, fbc)
DF_new <- rbind(DF_new, rgse)
DF_new <- rbind(DF_new, bac)
DF_new <- rbind(DF_new, tvix)
DF_new <- rbind(DF_new, uvxy)

#check for NAs
summary(DF_new)

#View Class of each variable
sapply(DF_new, class)

#Find Stock with -1 Low
DF_new %>% filter(Low == -1)
#HLG on 2017-08-10, low should be 10.13, high should be 10.98,
#close should be 10.35 and Volume of 32,600

DF_new %>% filter(Symbol == "hlg") %>% filter(Date >= as.Date("2017-08-1") &
Date <= as.Date("2017-08-20"))

#Replace incorrect HLG data
DF_new$Low[DF_new$Low == "-1"] <- "10.13"
DF_new[DF_new$High == "863.473" &
DF_new$Symbol == "hlg", 4] <- "10.98"
DF_new[DF_new$Close == "863.473" &
DF_new$Symbol == "hlg", 6] <- "10.35"
DF_new[DF_new$Volume == "10" & DF_new$Symbol == "hlg", 7] <- "32600"



#change from scientific notation
options(scipen = 999)

#check summary
summary(DF_new)

#Find Stock with 1423712891 Open
DF_new %>% filter(Open == 1423712891) #drys checked on yahoo finance and correct
DF_new %>% filter(Symbol == "drys") %>% filter(Date >= as.Date("2007-10-10") &
Date <= as.Date("2007-10-20"))

#Find Stock with 2423735131 Volume
DF_new %>% filter(Volume == 2423735131)
#View days around that date
DF_new %>% filter(Symbol == "bac") %>% filter(Date >= as.Date("2012-03-01") &
Date <= as.Date("2012-03-10"))
#bac checked on yahoo finance and incorrect should be 328,331,900
#replace incorrect volume
DF_new[DF_new$Volume == "2423735131" &
DF_new$Symbol == "bac", 7] <- "328331900"

#Find Stock with 2304018600 Volume
DF_new %>% filter(Volume == 2304018600)  #brk_b
#View days around that date
DF_new %>% filter(Symbol == "brk_b") %>% filter(Date >= as.Date("2010-02-07") &
Date <= as.Date("2010-02-15"))  #appears correct based on morningstar
# http://performance.morningstar.com/stock/performance-return.action?p=price_history_page&t=BRK.B&region=usa&culture=en-US

#Change columns back to numeric
DF_new$Low <- as.numeric(DF_new$Low)
DF_new$High <- as.numeric(DF_new$High)
DF_new$Close <- as.numeric(DF_new$Close)
DF_new$Volume <- as.numeric(DF_new$Volume)

#check class
sapply(DF_new, class)

```
#Feature Creation  

##Log Prices and Volumes 

Lead: I'm going to convert stock and ETF prices to log prices and volume to log volume.   

```{r}
#Calculate log prices and volumes and create new columns
DF_new <- DF_new %>%
mutate(
Log_Open = log(Open),
Log_High = log(High),
Log_Low = log(Low),
Log_Close = log(Close),
Log_Volume = log(Volume)
) %>%
ungroup()

#graphically compare Arithmetic vs log prices/volumes
#load ggplot2
library(ggplot2)
#Opening Price Arithmetic
ggplot(data = DF_new, aes(DF_new$Open)) + geom_histogram() + xlab("Opening Price")

#Opening Price log
ggplot(data = DF_new, aes(DF_new$Log_Open)) + geom_histogram() + xlab("Log Opening Price")

#High Price Arithmetic
ggplot(data = DF_new, aes(DF_new$High)) + geom_histogram() + xlab("High Price")

#High Price log
ggplot(data = DF_new, aes(DF_new$Log_High)) + geom_histogram() + xlab("Log High Price")

#Low Price Arithmetic
ggplot(data = DF_new, aes(DF_new$Low)) + geom_histogram() + xlab("Low Price")

#Low Price log
ggplot(data = DF_new, aes(DF_new$Log_Low)) + geom_histogram() + xlab("Log Low Price")

#Closing Price Arithmetic
ggplot(data = DF_new, aes(DF_new$Close)) + geom_histogram() + xlab("Closinging Price")

#Closing Price log
ggplot(data = DF_new, aes(DF_new$Log_Close)) + geom_histogram() + xlab("Log Closing Price")

#Volume Arithmetic
ggplot(data = DF_new, aes(DF_new$Volume)) + geom_histogram() + xlab("Volume")

#Volume log
ggplot(data = DF_new, aes(DF_new$Log_Volume)) + geom_histogram() + xlab("Log Volume")
```
Analysis: As you can see in the above graphs converting to log values reduces the wide range of data into a more manageable size.  For example, with arithmetic scaling, all of the metrics were right skewed. For price, DRYS drove this due to the very high stock price.  For volume, brk_b drove this due to high volume.  Additionally, the log graphs, are log-normally distributed.

Conclusion: Converting price and volume to log scale enabled better visualization of the data.  Additionally, by converting to log scale the metrics approximately follow a normal distribution enabling the use of many statistical tests that require the assumption of normality.  (Source: http://www.automatedtrader.net/articles/strategies/158619/models-for-daily-and-intra_day-volume-prediction.) 

##Log Returns 

Lead: I'm going to calculate the daily log returns of each stock and ETF.   

```{r}
#Calculate daily log return, add new column for variable
DF_new$Log_Daily_Return <- log(DF_new$Close / DF_new$Open)
#summary
summary(DF_new)
#find INF results for log return
DF_new %>% filter(Log_Daily_Return == "Inf") #Log Daily Return infinite when opening price = 0

#check to see if data is correct
#bcom
DF_new %>% filter(Symbol == "bcom") %>% filter(Date >= as.Date("2011-03-10") &
Date <= as.Date("2011-03-20"))
#open and low price missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "bcom", 3] <- "19.301"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "bcom", 5] <- "19.301"

#blj
DF_new %>% filter(Symbol == "blj") %>% filter(Date >= as.Date("2008-02-02") &
Date <= as.Date("2008-02-10"))
#open, low volume missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "blj", 3] <- "12.138"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "blj", 5] <- "12.138"
DF_new[DF_new$Volume == "0" & DF_new$Symbol == "blj", 7] <- "1464"

#cwi
DF_new %>% filter(Symbol == "cwi") %>% filter(Date >= as.Date("2007-01-07") &
Date <= as.Date("2007-01-20"))
#open, low volume missing, impute with prior days close price
#didn't start trading until 1/17/2007
#source:https://finance.yahoo.com/quote/CWI/history?period1=1167886800&period2=1169269200&interval=1d&filter=history&frequency=1d
#remove rows 2007-01-10 to 2007-01-16
DF_new = DF_new[!(
DF_new$Date >= as.Date("2007-01-10") &
DF_new$Date <= as.Date("2007-01-16") & DF_new$Symbol == "cwi"
), ]

#drh
#didn't start trading until 2005-05-25
#https://finance.yahoo.com/quote/DRH/history?period1=1116561600&period2=1117425600&interval=1d&filter=history&frequency=1d
#remove row 2005-05-25
DF_new = DF_new[!(DF_new$Date == as.Date("2005-05-25") &
DF_new$Symbol == "drh"), ]

#efv
DF_new %>% filter(Symbol == "efv") %>% filter(Date >= as.Date("2005-10-06") &
Date <= as.Date("2005-10-15"))
#open, low volume missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "efv", 3] <- "41.368"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "efv", 5] <- "41.368"
DF_new[DF_new$Volume == "0" & DF_new$Symbol == "efv", 7] <- "10792"

#emi
DF_new %>% filter(Symbol == "emi") %>% filter(Date >= as.Date("2007-03-25") &
Date <= as.Date("2007-03-30"))
#open, low volume missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "emi", 3] <- "9.837"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "emi", 5] <- "9.837"
DF_new[DF_new$Volume == "0" & DF_new$Symbol == "emi", 7] <- "31695"

#emj
DF_new %>% filter(Symbol == "emj") %>% filter(Date >= as.Date("2008-02-01") &
Date <= as.Date("2008-02-10"))
#open, low volume missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "emj", 3] <- "10.308"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "emj", 5] <- "10.308"
DF_new[DF_new$Volume == "0" & DF_new$Symbol == "emj", 7] <- "2872"

#fmo
DF_new %>% filter(Symbol == "fmo") %>% filter(Date >= as.Date("2007-11-01") &
Date <= as.Date("2007-11-07"))
#open, low missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "fmo", 3] <- "13.618"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "fmo", 5] <- "13.618"

#fud
DF_new %>% filter(Symbol == "fud") %>% filter(Date >= as.Date("2008-03-28") &
Date <= as.Date("2008-05-01"))
#didn't start trading until 2008-04-03 source:
#https://finance.yahoo.com/quote/FUD/history?period1=1207022400&period2=1211688000&interval=1d&filter=history&frequency=1d

#remove row 2008-04-01
DF_new = DF_new[!(DF_new$Date == as.Date("2008-04-01") &
DF_new$Symbol == "fud"), ]

#gbb
DF_new %>% filter(Symbol == "gbb") %>% filter(Date >= as.Date("2007-05-05") &
Date <= as.Date("2007-05-10"))
#remove row 2007-05-08
DF_new = DF_new[!(DF_new$Date == as.Date("2007-05-08") &
DF_new$Symbol == "gbb"), ]

#ite
DF_new %>% filter(Symbol == "ite") %>% filter(Date >= as.Date("2007-05-20") &
Date <= as.Date("2007-06-10"))
#remove rows 2007-01-10 to 2007-01-16
#didn't start trading
#https://finance.yahoo.com/quote/ITE/history?period1=1181448000&period2=1183176000&interval=1d&filter=history&frequency=1d
DF_new = DF_new[!(
DF_new$Date >= as.Date("2007-05-23") &
DF_new$Date <= as.Date("2007-05-29") & DF_new$Symbol == "ite"
), ]

#itm
DF_new %>% filter(Symbol == "itm") %>% filter(Date >= as.Date("2009-04-15") &
Date <= as.Date("2009-04-25"))
#open, low missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "itm", 3] <- "17.608"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "itm", 5] <- "17.608"

#mln
DF_new %>% filter(Symbol == "mln") %>% filter(Date >= as.Date("2009-04-15") &
Date <= as.Date("2009-04-30"))
#open, low missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "mln", 3] <- "13.397"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "mln", 5] <- "13.397"

#nom
DF_new %>% filter(Symbol == "nom") %>% filter(Date >= as.Date("2008-02-02") &
Date <= as.Date("2008-02-10"))
#open, low volume missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "nom", 3] <- "11.246"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "nom", 5] <- "11.246"
DF_new[DF_new$Volume == "0" & DF_new$Symbol == "nom", 7] <- "100"

#smb
DF_new %>% filter(Symbol == "smb") %>% filter(Date >= as.Date("2009-04-15") &
Date <= as.Date("2009-04-25"))
#open, low missing, impute with prior days close price
DF_new[DF_new$Open == "0" & DF_new$Symbol == "smb", 3] <- "15.489"
DF_new[DF_new$Low == "0" & DF_new$Symbol == "smb", 5] <- "15.489"

#spab
DF_new %>% filter(Symbol == "spab") %>% filter(Date >= as.Date("2007-05-20") &
Date <= as.Date("2007-05-30"))
#remove rows 2007-05-23 to 2007-05-29
DF_new = DF_new[!(
DF_new$Date >= as.Date("2007-05-23") &
DF_new$Date <= as.Date("2007-05-29") & DF_new$Symbol == "spab"
), ]

#sptl
DF_new %>% filter(Symbol == "sptl") %>% filter(Date >= as.Date("2007-05-20") &
Date <= as.Date("2007-05-30"))
#remove rows 2007-05-23 to 2007-05-29
DF_new = DF_new[!(
DF_new$Date >= as.Date("2007-05-23") &
DF_new$Date <= as.Date("2007-05-29") & DF_new$Symbol == "sptl"
), ]

#uci
DF_new %>% filter(Symbol == "uci") %>% filter(Date >= as.Date("2008-03-25") &
Date <= as.Date("2008-04-05"))
#didn't start trading until 2008-04-02
#https://finance.yahoo.com/quote/UCI/history?period1=1205553600&period2=1207195200&interval=1d&filter=history&frequency=1d
#remove row 2008-04-01
DF_new = DF_new[!(DF_new$Date == as.Date("2008-04-01") &
DF_new$Symbol == "uci"), ]

#usv
DF_new %>% filter(Symbol == "usv") %>% filter(Date >= as.Date("2008-03-25") &
Date <= as.Date("2008-04-05"))
#remove row 2008-04-01
DF_new = DF_new[!(DF_new$Date == as.Date("2008-04-01") &
DF_new$Symbol == "usv"), ]

#vxz
DF_new %>% filter(Symbol == "vxz") %>% filter(Date >= as.Date("2009-01-28") &
Date <= as.Date("2009-02-05"))
#remove row 2009-01-29
DF_new = DF_new[!(DF_new$Date == as.Date("2009-01-29") &
DF_new$Symbol == "vxz"), ]

#rwx
#didn't start trading until 12/19/2006
#source: http://www.kibot.com/Historical_Data/All_Stocks_And_ETFs_Historical_Intraday_Data.aspx
#remove rows 2006-12-15 to 2006-12-18
DF_new = DF_new[!(
DF_new$Date >= as.Date("2006-12-15") &
DF_new$Date <= as.Date("2006-12-18") & DF_new$Symbol == "rwx"
), ]

#summary
summary(DF_new)

#Check very large returns to make sure they are correct
#filter for log price changes greater than or less than 4%
DF_change <- DF_new %>% filter(!between(Log_Daily_Return, -4, 4))
View(DF_change)

#ako-a
#View days around that date
DF_new %>% filter(Symbol == "ako-a") %>% filter(Date >= as.Date("2010-08-25") &
Date <= as.Date("2010-09-04")) #incorrect open and low prices
#open, low wrong, impute with prior days close price
DF_new[DF_new$Open == "0.00924" &
DF_new$Symbol == "ako-a", 3] <- "21.622"
DF_new[DF_new$Low == "0.00924" &
DF_new$Symbol == "ako-a", 5] <- "21.622"

#avb
DF_new %>% filter(Symbol == "avb") %>% filter(Date >= as.Date("2010-09-12") &
Date <= as.Date("2010-09-19"))
#closing price wrong, change to 87 (source: yahoo finance)
DF_new[DF_new$Close == "0.8997" & DF_new$Symbol == "avb", 6] <- "87"

#bofil
DF_new %>% filter(Symbol == "bofil") %>% filter(Date >= as.Date("2016-03-07") &
Date <= as.Date("2016-03-12"))
#open, high wrong, impute with prior days close price
DF_new[DF_new$Open == "182681.41" &
DF_new$Symbol == "bofil", 3] <- "22.516"
DF_new[DF_new$High == "182681.41" &
DF_new$Symbol == "bofil", 4] <- "22.516"

#bxp
DF_new %>% filter(Symbol == "bxp") %>% filter(Date >= as.Date("2007-04-30") &
Date <= as.Date("2007-05-10"))
#close wrong, should be 87.88 (source: yahoo finance)
DF_new[DF_new$Close == "0.8874" &
DF_new$Symbol == "bxp", 6] <- "87.88"
#open wrong, should be 87.35 (source: yahoo finance)
DF_new[DF_new$Open == "0.8874" &
DF_new$Symbol == "bxp", 3] <- "87.35"

#cbmxw
DF_new %>% filter(Symbol == "cbmxw") %>% filter(Date >= as.Date("2016-05-28") &
Date <= as.Date("2016-06-3")) #correct

#fud
DF_new %>% filter(Symbol == "fud") %>% filter(Date >= as.Date("2008-10-06") &
Date <= as.Date("2008-10-11"))
#open, low wrong, fix source: yahoo finance
DF_new[DF_new$Open == "0.3" & DF_new$Symbol == "fud", 3] <- "17.65"
DF_new[DF_new$Low == "0.3" & DF_new$Symbol == "fud", 5] <- "17.39"

#gcbc
DF_new %>% filter(Symbol == "gcbc") %>% filter(Date >= as.Date("2016-04-25") &
Date <= as.Date("2016-04-30"))
#open, high wrong source:yahoo finance
DF_new[DF_new$Open == "77739.67" &
DF_new$Symbol == "gcbc", 3] <- "17.81"
DF_new[DF_new$High == "77739.67" &
DF_new$Symbol == "gcbc", 4] <- "19"

#phii
DF_new %>% filter(Symbol == "phii") %>% filter(Date >= as.Date("2016-04-12") &
Date <= as.Date("2016-04-22"))

#open, high wrong source:yahoo finance
DF_new[DF_new$Open == "76800.02" &
DF_new$Symbol == "phii", 3] <- "18.90"
DF_new[DF_new$High == "76800.02" &
DF_new$Symbol == "phii", 4] <- "20.95"

#slg
DF_new %>% filter(Symbol == "slg") %>% filter(Date >= as.Date("2014-03-01") &
Date <= as.Date("2014-03-07"))
#closing price wrong  (source: yahoo finance)
DF_new[DF_new$Close == "0.9161" &
DF_new$Symbol == "slg", 6] <- "90.11"

#smbk
DF_new %>% filter(Symbol == "smbk") %>% filter(Date >= as.Date("2016-01-10") &
Date <= as.Date("2016-01-20"))
#open, high wrong source:yahoo finance
DF_new[DF_new$Open == "199999.99" &
DF_new$Symbol == "smbk", 3] <- "16.90"
DF_new[DF_new$High == "199999.99" &
DF_new$Symbol == "smbk", 4] <- "16.90"

#smed
DF_new %>% filter(Symbol == "smed") %>% filter(Date >= as.Date("2016-07-20") &
Date <= as.Date("2016-07-25"))
#open, low wrong, fix source: yahoo finance
DF_new[DF_new$Open == "0.0096" &
DF_new$Symbol == "smed", 3] <- "5.38"
DF_new[DF_new$Low == "0.0093" & DF_new$Symbol == "smed", 5] <- "5"

#tgen
DF_new %>% filter(Symbol == "tgen") %>% filter(Date >= as.Date("2016-03-15") &
Date <= as.Date("2016-03-25"))
#open, high wrong source:yahoo finance
DF_new[DF_new$Open == "80000.99" &
DF_new$Symbol == "tgen", 3] <- "3.85"
DF_new[DF_new$High == "80000.99" &
DF_new$Symbol == "tgen", 4] <- "3.86"

#wsr
DF_new %>% filter(Symbol == "wsr") %>% filter(Date >= as.Date("2010-09-28") &
Date <= as.Date("2010-10-15"))
#open, high wrong source:yahoo finance
DF_new[DF_new$Open == "1133.16" &
DF_new$Symbol == "wsr", 3] <- "6.25"
DF_new[DF_new$High == "1133.16" &
DF_new$Symbol == "wsr", 4] <- "6.25"
DF_new[DF_new$Open == "570.08" &
DF_new$Symbol == "wsr", 3] <- "6.70"
DF_new[DF_new$High == "570.08" &
DF_new$Symbol == "wsr", 4] <- "6.70"


#Change columns back to numeric
DF_new$Low <- as.numeric(DF_new$Low)
DF_new$Open <- as.numeric(DF_new$Open)
DF_new$High <- as.numeric(DF_new$High)
DF_new$Close <- as.numeric(DF_new$Close)
DF_new$Volume <- as.numeric(DF_new$Volume)

#check class
sapply(DF_new, class)

#rerun daily log return
DF_new$Log_Daily_Return <- log(DF_new$Close / DF_new$Open)

#summary, check to see if INF and large daily changes are gone
summary(DF_new)

#filter for log price changes greater than or less than 3%
DF_change2 <- DF_new %>% filter(!between(Log_Daily_Return, -3, 3))
#cbmxw only stock with log daily return/loss of greater than 4%
#due to time constraints I did not validate all of the stock and ETF prices, though I expect some of the stocks/ETFs showing prices changes of greater than log 3% are likely incorrect.

#graph
DF_change2$Symbol <- as.factor(DF_change2$Symbol)
ggplot(DF_change2,
aes(
x = Date,
y = Log_Daily_Return,
position = "stack",
fill = Symbol
)) +
geom_bar(stat = "identity", width = 5) + geom_text(data = subset(DF_change2, Symbol == 'cbmxw' |
Symbol == 'syn'),
aes(label = Symbol))

#Daily return Arithmetic
#Calculate daily return arithmetic
DF_new$Daily_Return <- ((DF_new$Close / DF_new$Open) - 1)
#graph daily return
ggplot(data = DF_new, aes(DF_new$Daily_Return)) + geom_histogram() + xlab("Daily Return")

#graph Daily return log
ggplot(data = DF_new, aes(DF_new$Log_Daily_Return)) + geom_histogram() + xlab("Log Daily Return")

```
  
Analysis: Log return equaled infinite when opening price was $0.  This occurred due to errors in the data, either the data was incorrect (i.e. wrong price) or the stock/ETF had not started trading as of that date (i.e. prior to IPO). When the data was incorrect, I fixed the data.  When the stock had not started trading as of that date, I removed the row from the data frame. Additionally, the log return variable revealed incorrect data.  I saw large changes in price.  Upon further inspection, many of the data inputs were incorrect.  I corrected the prices.  This resulted in one stock showing a significant daily change, CBMXW. It had a log daily return of 4.335507% on 2016-05-31. As you can see in the above graph, SYN had the greatest frequency of large stock moves (7 days) and CBMXW had the greatest daily change in price.  

Conclusion: By calculating this variable, I was able to discover errors in the data that required fixing. I was also able to see which stocks/ETF's had the greatest daily price change, and which stocks/ETF's had the largest volatility.  
Additionally, the benefit of using log returns instead of absolute prices is normalization, "measuring all variables in a comparable metric, thus enabling evaluation of analytic relationships among two or more variables despite originating from price series of unequal values." (source: https://quantivity.wordpress.com/2011/02/21/why-log-returns/). Another benefit of log returns is, assuming normal distribution, adding period returns produces an end period return that is also normally distributed (source: https://www.youtube.com/watch?v=PtoUlt3V0CI). Additionally,graphing log returns enables a person to see big moves on a percentage basis not on an absolute dollar basis (source: https://www.usatoday.com/story/money/columnist/krantz/2013/08/25/linear-logarithmic-stock-charts/2657493/). Also, converting prices to a log scale is highly useful when charting stock prices, as a significant percentage move will always correspond to a significant visual change (source :https://finance.zacks.com/use-logarithmic-scale-stocks-8760.html)   

##Rolling moving average   

Lead: I'm going to calculate the 5 day moving average on the closing price of each stock and ETF.   

```{r}
#load library
library(RcppRoll)
#caclulate 200 day moving average on closing price, create new column
DF_new <- DF_new %>%
group_by(Symbol) %>%
mutate(Roll_Avg_Price = roll_mean(
Close,
200,
na.rm = TRUE,
align = "right",
fill = 0
)) %>%
ungroup()

#summary
summary(DF_new)

#graph rolling average vs. closing price for a select stock, GS
ggplot() +
geom_line(data = subset(DF_new, Symbol == 'gs'),
aes(x = Date, y = Close, colour = "blue")) +
geom_line(data = subset(DF_new, Symbol == 'gs'),
aes(x = Date, y = Roll_Avg_Price, colour = "red")) +
labs(
title = "Price vs. Moving Avg",
x = "Date",
y = "Price",
color = "Legend"
) +
scale_color_manual(
labels = c("Closing Price", "200 Day Moving Avg"),
values = c("blue", "red")
)

```
Analysis: As you can see in the above graph, the 200 day moving average smooths out the data.  
A moving average can also act as support or resistance. In a downtrend, a moving average may act as resistance; like a ceiling, the price hits the level and then starts to drop again.  This can be seen in the GS graph during 2008.  

Conclusion: The moving average is a common technical analysis tool used by traders.  It is often used to identify trend direction and can also be used to generate potential buy and sell signals (source: http://www.onlinetradingconcepts.com/TechnicalAnalysis/MASimple.html).  

##Feature Creation: Day of Week  

Lead: Add day of week to the data frame.   

```{r}
#add day of week to data frame
DF_new$Day <- weekdays(DF_new$Date)

#calculate return by day of week for each ETF and stocks
DF_new <- DF_new %>%
  group_by(Symbol, Day) %>%
  mutate(Weekday_Return = mean(Log_Daily_Return)) %>%
  ungroup()

#summary
summary(DF_new)

#filter
DF_new %>% filter(Symbol == "gs") %>% filter(Date >= as.Date("1999-05-10") &
                                               Date <= as.Date("1999-05-14")) # Wednesday's achieved the highest returns

#calculate return by day of week for all ETF and stocks
DF_new <- DF_new %>%
  group_by(Day) %>%
  mutate(Weekday_Return_All = mean(Log_Daily_Return)) %>%
  ungroup()

#summary
summary(DF_new)

#filter
DF_new %>% filter(Symbol == "gs") %>% filter(Date >= as.Date("1999-05-10") &
                                               Date <= as.Date("1999-05-14")) # Friday's achieved the highest returns

#graph
#filter
DF_days = DF_new %>% filter(Symbol == "gs") %>% filter(Date >= as.Date("1999-05-10") &
                                                         Date <= as.Date("1999-05-14"))
#set factor levels
DF_days$Day <- as.factor(DF_days$Day)
DF_days$Day <-
  factor(DF_days$Day,
         levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))

#graph for all stocks and etfs combined
ggplot(DF_days, aes(Day, Weekday_Return_All)) +
  geom_col() + ylab("Daily Log Return") + ggtitle("All Stocks and ETFs")

#graph for GS
ggplot(DF_days, aes(Day, Weekday_Return)) +
  geom_col() + ylab("Daily Log Return") + ggtitle("GS")
```
Analysis: When looking at all stocks and ETF's combined, Friday's achieved the highest daily log return, Monday's performed worst.  Looking at a specific stock, GS, Wednesday's performed best while Monday's again performed worst.

Conclusion: There seems to be a bias towards positive market performance on Friday's and a bias towards under performance on Monday's.  The business insight for this is investors are better off buying on Monday's selling on Friday's. Additionally, this analysis can be done on specific stocks. For example, over GS's trading history, Wednesday's outperformed and Monday's performed worst.  
